# H2O GBM Tuning Tutorial for R
## Arno Candel, PhD, Chief Architect, H2O.ai

In this tutorial, we show how to build a well-tuned GBM model for a supervised classification task.
This tutorial is written in Rmarkdown. You can download the source [here](https://github.com/h2oai/h2o-3/blob/master/h2o-docs/src/product/tutorials/gbm/gbmTuning.Rmd).

## Installation of the H2O R Package
Either download H2O from [H2O.ai's website](http://h2o.ai/download) or install the latest version of H2O into R with the following R code:

```{r message=FALSE, warning=FALSE}
# The following two commands remove any previously installed H2O packages for R.
if ("package:h2o" %in% search()) { detach("package:h2o", unload=TRUE) }
if ("h2o" %in% rownames(installed.packages())) { remove.packages("h2o") }

# Next, we download packages that H2O depends on.
pkgs <- c("methods","statmod","stats","graphics","RCurl","jsonlite","tools","utils")
for (pkg in pkgs) {
  if (! (pkg %in% rownames(installed.packages()))) { install.packages(pkg) }
}

# Now we download, install and initialize the H2O package for R.
install.packages("h2o", type="source", repos=(c("http://h2o-release.s3.amazonaws.com/h2o/rel-turchin/6/R")))
```

## Launch an H2O cluster on localhost

```{r message=FALSE, warning=FALSE}
library(h2o)
h2o.init(nthreads=-1)
## optional: connect to a running H2O cluster
#h2o.init(ip="mycluster", port=55555) 
```

## Import the data into H2O 
Everything is scalable and distributed from now on. All processing is done on the fully multi-threaded and distributed H2O Java-based backend and can be scaled to large datasets on large compute clusters.
Here, we use a small public dataset [Titanic](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/Titanic.html), but you can use datasets that are hundreds of GBs large.

```{r message=FALSE, warning=FALSE}
## 'path' can point to a local file, hdfs, s3, nfs, Hive, directories, etc.
df <- h2o.importFile(path = "http://s3.amazonaws.com/h2o-public-test-data/smalldata/gbm_test/titanic.csv")
dim(df)
head(df)
tail(df)
summary(df,exact_quantiles=TRUE)

## pick a response for the supervised problem
response <- "survived"

## the response variable is an integer, we will turn it into a categorical/factor for binary classification
df[[response]] <- as.factor(df[[response]])           

## use all other columns (except for the name) as predictors
predictors <- setdiff(names(df), c(response, "name")) 
```

From now on, everything is generic and directly applies to most datasets.

## Split the data for Machine Learning
We split the data into three pieces: 60% for training, 20% for validation, 20% for final testing. 
Here, we use random splitting, but this assumes i.i.d. data. If this is not the case (e.g., when events span across multiple rows or data has a time structure), you'll have to sample your data non-randomly.

```{r message=FALSE, warning=FALSE}
splits <- h2o.splitFrame(
  data = df, 
  ratios = c(0.6,0.2),   ## only need to specify 2 fractions, the 3rd is implied
  destination_frames = c("train.hex", "valid.hex", "test.hex"), seed = 1234
)
train <- splits[[1]]
valid <- splits[[2]]
test  <- splits[[3]]
```

## Establish baseline performance
As the first step, we'll build some default models to see what accuracy we can expect. Let's use the [AUC metric](http://mlwiki.org/index.php/ROC_Analysis) for this demo, but you can use `h2o.logloss` and `stopping_metric="logloss"` as well.


The first model is a default GBM, trained on the 60% training split

```{r message=FALSE, warning=FALSE}
gbm <- h2o.gbm(x = predictors, y = response, training_frame = train)

## Show a detailed model summary
gbm

## Get the AUC on the validation set
h2o.auc(h2o.performance(gbm, newdata = valid)) 
```


The second model is a default GBM, trained on 80% of the data (here, we combine the training and validation splits to get more training data), and cross-validated using 4 folds.
Note that cross-validation takes longer and is not usually done for really large datasets.

```{r message=FALSE, warning=FALSE}
## h2o.rbind makes a copy here, so it's better to use splitFrame with ratios = 0.8 instead if you want 
gbm <- h2o.gbm(x = predictors, y = response, training_frame = h2o.rbind(train, valid), nfolds = 4, seed = 0xDECAF)

## Show a detailed summary of the cross validation metrics
## This gives you an idea of the variance between the folds
gbm@model$cross_validation_metrics_summary

## Get the cross-validated AUC by scoring the combined holdout predictions.
## (Instead of taking the average of the metrics across the folds)
h2o.auc(h2o.performance(gbm, xval = TRUE))
```


Next, we train a GBM with "I feel lucky" parameters (which usually are better than defaults for realistic large datasets).
We'll use early stopping to automatically tune the number of trees using the validation AUC. 
We'll use a lower learning rate (lower is always better, just takes more trees to converge).
We'll also use stochastic sampling of rows and columns to (hopefully) improve generalization.

```{r message=FALSE, warning=FALSE}
gbm <- h2o.gbm(
  ## standard model parameters
  x = predictors, 
  y = response, 
  training_frame = train, 
  validation_frame = valid,
  
  ## more trees is better if the learning rate is small enough 
  ## here, use "more than enough" trees - we have early stopping
  ntrees = 10000,                                                            
  
  ## smaller learning rate is better (this is a good value for most datasets, but see below for annealing)
  learn_rate=0.01,                                                         
  
  ## early stopping once the validation AUC doesn't improve by at least 0.01% for 5 consecutive scoring events
  stopping_rounds = 5, stopping_tolerance = 1e-4, stopping_metric = "AUC", 
  
  ## sample 80% of rows per tree
  sample_rate = 0.8,                                                       
  
  ## sample 80% of columns per split
  col_sample_rate = 0.8,                                                   
  
  ## fix a random number generator seed for reproducibility
  seed = 1234,                                                             
  
  ## score every 10 trees to make early stopping reproducible (it depends on the scoring interval)
  score_tree_interval = 10                                                 
)

## Get the AUC on the validation set
h2o.auc(h2o.performance(gbm, valid = TRUE))
```


Hint: Specify an invalid stopping_metric to see what other choices exist:

```{r message=FALSE, warning=FALSE}
gbm <- h2o.gbm(x = predictors, y = response, training_frame = train, stopping_metric = "yada")
```

## Hyper-Parameter Search

Next, we'll do real hyper-parameter optimization to see if we can beat the best AUC so far (around 94%).

First we want to know what value of `max_depth` to use (this is dataset specific, and has a big impact on runtime).
We'll do a quick Cartesian grid search to get a rough idea. Each model in the grid search will use early stopping to tune the number of trees using the validation set AUC, as before.
We'll use learning rate annealing to speed up convergence without sacrificing too much accuracy.

```{r message=FALSE, warning=FALSE}
## try a broad range for max_depth (depth 20 is usually too much and takes a lot longer than 5 or 10)
## for some datasets, however, you might need even more depth
hyper_params = list( max_depth = c(1:20) ) 

grid <- h2o.grid(
  ## hyper parameters
  hyper_params = hyper_params,
  
  ## full Cartesian hyper-parameter search
  search_criteria = list(strategy = "Cartesian"),
  
  ## which algorithm to run
  algorithm="gbm",
  
  ## identifier for the grid, to later retrieve it
  grid_id="depth_grid",
  
  ## standard model parameters
  x = predictors, 
  y = response, 
  training_frame = train, 
  validation_frame = valid,
  
  ## more trees is better if the learning rate is small enough 
  ## here, use "more than enough" trees - we have early stopping
  ntrees = 10000,                                                            
  
  ## smaller learning rate is better
  ## since we have learning_rate_annealing, we can afford to start with a bigger learning rate
  learn_rate = 0.05,                                                         
  
  ## learning rate annealing: learning_rate shrinks by 1% after every tree 
  ## (use 1.00 to disable, but then lower the learning_rate)
  learn_rate_annealing = 0.99,                                               
  
  ## sample 80% of rows per tree
  sample_rate = 0.8,                                                       
  
  ## sample 80% of columns per split
  col_sample_rate = 0.8,                                                   
  
  ## fix a random number generator seed for reproducibility
  seed = 1234,                                                             
  
  ## early stopping once the validation AUC doesn't improve by at least 0.01% for 5 consecutive scoring events
  stopping_rounds = 5,
  stopping_tolerance = 1e-4,
  stopping_metric = "AUC", 
  
  ## score every 10 trees to make early stopping reproducible (it depends on the scoring interval)
  score_tree_interval = 10                                                
)

## by default, display the grid search results sorted by logloss (since this is a classification task)
grid                                                                       

## sort the grid models by AUC
sortedGrid <- h2o.getGrid("depth_grid", sort_by="auc", decreasing = TRUE)    
sortedGrid

## find the range of max_depth for the top 10 models
top10Depths = sortedGrid@summary_table$max_depth[1:10]                       
minDepth = min(as.numeric(top10Depths))
maxDepth = max(as.numeric(top10Depths))
```

Now that we know a good range for max_depth, we can tune all other parameters in more detail
We'll use random hyper-parameter search to "let the machine get luckier than the best guess of a human".

```{r message=FALSE, warning=FALSE}
hyper_params = list( 
  ## restrict the search to the range of max_depth established above
  max_depth = seq(minDepth,maxDepth,1),                                      
  
  ## search a large space of row sampling rates per tree
  sample_rate = seq(0.3,1,0.01),                                             
  
  ## search a large space of column sampling rates per split
  col_sample_rate = seq(0.3,1,0.01),                                         
  
  ## search a large space of column sampling rates per tree
  col_sample_rate_per_tree = seq(0.3,1,0.01),                                
  
  ## search a large space of how column sampling per split should change as a function of the depth of the split
  col_sample_rate_change_per_level = seq(0.9,1.1,0.01),                      
  
  ## search a large space of the number of min rows in a terminal node
  min_rows = 2^seq(0,log2(nrow(train))-1,1),                                 
  
  ## search a large space of the number of bins for split-finding for continuous and integer columns
  nbins = 2^seq(4,10,1),                                                     
  
  ## search a large space of the number of bins for split-finding for categorical columns
  nbins_cats = 2^seq(4,12,1),                                                
  
  ## search a few minimum required relative error improvement thresholds for a split to happen
  min_split_improvement = c(0,1e-8,1e-6,1e-4),                               
  
  ## try all histogram types (can be useful if there are numeric columns with extreme outliers)
  histogram_type = c("UniformAdaptive","QuantilesGlobal","RoundRobin")       
)

search_criteria = list(
  ## Random grid search
  strategy = "RandomDiscrete",      
  
  ## limit the max runtime to 10 hours
  max_runtime_secs = 36000,         
  
  ## build no more than 50 models (usually want a little more than that)
  max_models = 50,                  
  
  ## random number generator seed to make sampling of parameter combinations reproducible
  seed = 1234,                        
  
  ## early stopping once the leaderboard of the top 5 models is converged
  stopping_rounds = 5,                
  stopping_metric = "AUC",
  stopping_tolerance = 1e-4
)

grid <- h2o.grid(
  ## hyper parameters
  hyper_params = hyper_params,
  
  ## hyper-parameter search configuration (see above)
  search_criteria = search_criteria,
  
  ## which algorithm to run
  algorithm = "gbm",
  
  ## identifier for the grid, to later retrieve it
  grid_id = "final_grid", 
  
  ## standard model parameters
  x = predictors, 
  y = response, 
  training_frame = train, 
  validation_frame = valid,
  
  ## more trees is better if the learning rate is small enough
  ## use "more than enough" trees - we have early stopping
  ntrees = 10000,                                                            
  
  ## smaller learning rate is better
  ## since we have learning_rate_annealing, we can afford to start with a bigger learning rate
  learn_rate = 0.05,                                                         
  
  ## learning rate annealing: learning_rate shrinks by 1% after every tree 
  ## (use 1.00 to disable, but then lower the learning_rate)
  learn_rate_annealing = 0.99,                                               
  
  ## no model should take more than 1 hour, otherwise stop it early
  max_runtime_secs = 3600,                                                 
  
  ## early stopping once the validation AUC doesn't improve by at least 0.01% for 5 consecutive scoring events
  stopping_rounds = 5, stopping_tolerance = 1e-4, stopping_metric = "AUC", 
  
  ## score every 10 trees to make early stopping reproducible (it depends on the scoring interval)
  score_tree_interval = 10,                                                
  
  ## base random number generator seed for each model (automatically gets incremented internally for each model)
  seed = 1234                                                             
)

## Sort the grid models by AUC
sortedGrid <- h2o.getGrid("final_grid", sort_by = "auc", decreasing = TRUE)    
sortedGrid
```

We can see that the best models have even better validation AUCs than our previous best models, so the random grid search was successful!

```{r message=FALSE, warning=FALSE}
## Inspect the best 5 models from the grid search
for (i in 1:5) {
  gbm <- h2o.getModel(sortedGrid@model_ids[[i]])
  print(h2o.auc(h2o.performance(gbm, valid = TRUE)))
}

## Final (one-time) testing of the best model on the held out test set
gbm <- h2o.getModel(sortedGrid@model_ids[[1]])
print(h2o.auc(h2o.performance(gbm, newdata = test)))

## Make test set predictions
preds <- h2o.predict(gbm, test)

## Save the test set predictions to a file
#h2o.exportFile(preds, "/tmp/preds.csv", force=TRUE)
```